---
title: "PDO-NPGO GLMM and GAMM"
author: "Dr. Curry J. Cunningham"
date: "2/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

#Purpose
The purpose of this analysis is to use generalized additive mixed-effects models **GAMMs** to describe how the relationship between **PDO** and **NPGO**, and the productivity of Alaska, British Columbia, Washington, Oregon, and California salmon has **changed** over time. 

#Methods
##Load required packages
```{r}
require(tidyverse)
require(mgcv)
require(gamm4)
require(lme4)
require(kableExtra)
require(MuMIn)
require(coefplot2)
```

##Load and filter SR data
```{r}
dat <- read.csv(file.path("data","AK-WCoast-Salmon-SR.csv"), header=TRUE, stringsAsFactors=TRUE)
#Add rps and ln.rps
dat$rps <- dat$rec/dat$spawn
dat$ln.rps <- log(dat$rps)
#Filter for nonsense recruits-per-spawner values and limit year range.
dat.2 <- dat %>% filter(!is.infinite(ln.rps), !is.na(ln.rps), broodYr>=1950, broodYr<=2010)
```

##Bind with NPGO and PDO data
First we need to define the species-specific offsets from the brood year for covariate interactions. 

The current species are: `r unique(dat.2$species)`.

```{r}
species <- unique(dat.2$species)
offsets <- data.frame(species,c(2,2,1,2,3))
names(offsets) <- c("species","offset")
#Print Out
kable(head(offsets), align="crr") %>% #, digits=c(3,1,1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
covar.dat <- read.csv(file.path("data","Covariates","covar.dat.csv"))
```

Lets loop through the data and combine the correct covariate data

```{r}
dat.2$PDO <- NA
dat.2$NPGO <- NA

i<- 1
for(i in 1:nrow(dat.2)) {
  temp.year <- dat.2$broodYr[i]
  covar.year <- temp.year + offsets$offset[offsets$species==dat.2$species[i]]
  #Add Covariate value for the appropriate offset year
  dat.2$PDO[i] <- covar.dat$avg[covar.dat$name=='pdo' & covar.dat$Year==covar.year] #PDO
  dat.2$NPGO[i] <- covar.dat$avg[covar.dat$name=='npgo' & covar.dat$Year==covar.year] #PDO
} #next i
```

Lets also add a factor for before/after the 1989 regime shift
```{r}
dat.2$era <- ifelse(dat.2$broodYr<1990, "pre", "post")


```


##Fit GLMMs
We will fit a series of generalized linear mixed models with a fixed breakpoint before (<=) and after **1989**. 

###Results

##Fit GAMMs
We will fit variable coefficient GAMs describing changes in the effect of **PDO** and **NPGO** over time. 

###Sockeye

```{r}
temp.dat <- dat.2 %>% filter(species=="Sockeye")
```

Lets check and see if there is evidence for a single **PDO** effect or assuming this effect varies randomly among stocks. 

```{r}
#Common PDO effect across stocks
m <- lme(ln.rps ~ 1 + stock:spawn + PDO, random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")

m.5 <- lme(ln.rps ~ 1 + PDO:region, random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")
m.51 <- lme(ln.rps ~ 1 + PDO + PDO:region, random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")
m.52 <- lme(ln.rps ~ 1 + region + PDO + PDO:region, random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")
m.53 <- lme(ln.rps ~ 1 + region + PDO:region, random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")
m.54 <- lme(ln.rps ~ 1 + PDO, random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")
AICc(m.5,m.51,m.52,m.53)

#Treating PDO Effect as Random Among Populations - Fixed across time
m1 <- lme(ln.rps ~ 1 + stock:spawn + PDO, random= ~1 + PDO | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="REML")

AIC <- AICc(m, m1) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Yes, it seems that there is evidence for a randomly varying **PDO** effect. 

Lets check the same thing with **NPGO** effect
```{r}
#Common NPGO effect across stocks
m <- lme(ln.rps ~ 1 + stock:spawn + NPGO, random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="REML")

#Treating NPGO Effect as Random Among Populations - Fixed across time
m1 <- lme(ln.rps ~ 1 + stock:spawn + NPGO, random= ~1 + NPGO | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="REML")

AIC <- AICc(m, m1) 
AIC <- AIC %>%
  mutate(model=rownames(AIC)) %>%
  mutate(dAICc=AICc-min(AICc)) %>%
  arrange(AICc)
AIC
```

Interestingly, there is a model-based preference for a common **NPGO** effect.

But, on to the main question: **Has the effect of NPGO and PDO varied over time?**

```{r}
g <- gamm(ln.rps ~ 1 + stock:spawn + PDO, random=list(stock=~1), 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
              control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")

g1 <- gamm(ln.rps ~ 1 + stock:spawn + PDO:region, random=list(stock=~1),
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
              control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")

g2 <- gamm(ln.rps ~ 1 + stock:spawn + s(broodYr, bs='cr', by=I(PDO:region)), random=list(stock=~1), 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
              control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 500, msMaxIter = 500), method="ML")
plot(g2$gam)


g3 <- gamm(ln.rps ~ 1 + stock:spawn + s(broodYr, bs='cr', by=(PDO:region)), random=list(stock=~1), 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
              control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")



AICc(g,g1,g2)

summary(g$gam)



```

###Chinook

###Pink

###Chum

###Coho

###Results

#General Conclusions

