---
title: "Testing for Breakpoint in PDO and NPGO Relationships"
author: "Dr. Curry J. Cunningham"
date: "2/5/2019"
output:
  html_document:
    collapsed: no
    fig_caption: yes
    highlight: espresso
    number_sections: yes
    smooth_scroll: yes
    theme: sandstone
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

#Purpose
The purpose of this analysis is to use generalized additive mixed-effects models **GAMMs** to describe how the relationship between **PDO** and **NPGO**, and the productivity of Alaska, British Columbia, Washington, Oregon, and California salmon has **changed** over time. 

#Methods
##Load required packages
```{r}
require(tidyverse)
require(mgcv)
require(gamm4)
require(lme4)
require(kableExtra)
require(MuMIn)
require(coefplot2)
```

##Load and filter SR data
```{r}
dat <- read.csv(file.path("data","AK-WCoast-Salmon-SR.csv"), header=TRUE, stringsAsFactors=TRUE)
#Add rps and ln.rps
dat$rps <- dat$rec/dat$spawn
dat$ln.rps <- log(dat$rps)
#Filter for nonsense recruits-per-spawner values and limit year range.
dat.2 <- dat %>% filter(!is.infinite(ln.rps), !is.na(ln.rps), broodYr>=1950, broodYr<=2010)
```

##Bind with NPGO and PDO data
First we need to define the species-specific offsets from the brood year for covariate interactions. 

The current species are: `r unique(dat.2$species)`.

```{r}
species <- unique(dat.2$species)
offsets <- data.frame(species,c(2,2,1,2,3))
names(offsets) <- c("species","offset")
#Print Out
kable(head(offsets), align="crr") %>% #, digits=c(3,1,1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
covar.dat <- read.csv(file.path("data","Covariates","covar.dat.csv"))
```

Lets loop through the data and combine the correct covariate data

```{r}
dat.2$PDO <- NA
dat.2$NPGO <- NA

i<- 1
for(i in 1:nrow(dat.2)) {
  temp.year <- dat.2$broodYr[i]
  covar.year <- temp.year + offsets$offset[offsets$species==dat.2$species[i]]
  #Add Covariate value for the appropriate offset year
  dat.2$PDO[i] <- covar.dat$avg[covar.dat$name=='pdo' & covar.dat$Year==covar.year] #PDO
  dat.2$NPGO[i] <- covar.dat$avg[covar.dat$name=='npgo' & covar.dat$Year==covar.year] #PDO
} #next i
```

Lets also add a factor for before/after the 1989 regime shift
```{r}
dat.2$era <- ifelse(dat.2$broodYr<1990, "pre", "post")


```


##Fit GLMMs
We will fit a series of generalized linear mixed models with a fixed breakpoint before (<=) and after **1989**. 

We will assume:
* A stock-specific autocorrelated error structure.
* Random effects structure where the intercept varies as a function of stock. 

The full model formula is:
```{r}
full.form <- formula(ln.rps ~ 1 + stock:spawn + era + PDO:large.region + NPGO:large.region +
                                                      PDO:large.region:era + NPGO:large.region:era)
```

###Sockeye

```{r}
temp.species <- "Sockeye"
temp.dat <- dat.2 %>% filter(species==temp.species)
```

First we will fit the full model, with differences in **PDO** and **NPGO** by region and by region and era. 
```{r}
s.full <- lme(full.form, random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="ML")
```

First, lets test whether there is evidence for differences in **PDO** effects among eras, through a **likelihood ratio test** for the `PDO:large.region:era` term.

```{r}
s.1 <- update(s.full, fixed=.~.-PDO:large.region:era)
anova(s.full, s.1)
```
We *cannot* reject the Null hypothesis that **era** has *no* effect on the **NPGO** relationship with population productivity, so we will *exclude* `PDO:large.region:era` in the `r temp.species` model. 

Next, lets conduct a likelihood ratio test for the change in regional **NPGO** effects among **era**s.

```{r}
s.2 <- update(s.1, fixed=.~.-NPGO:large.region:era)
anova(s.1, s.2)
```

Conversely, we *can* reject the Null hypothesis that **era** has *no* effect on the **NPGO** relationship with population productivity, so we will *include* `NPGO:large.region:era` in the `r temp.species` model.

Finally, lets conduct a LRT for the **era** term
```{r}
s.3 <- update(s.1, fixed=.~.-era)
anova(s.1, s.3)
```
Conversely, we *can* reject the Null hypothesis that **era** has *no* main effect, so we will *include* `era` in the `r temp.species` model.

Now, lets estimate the final `r temp.species` model with **REML**, 
```{r}
s.final <- lme(formula(s.1), random= ~1 | stock, 
              correlation=corAR1(form= ~ 1 | stock), data=temp.dat,
           control=lmeControl(opt="optim", optimMethod="Nelder-Mead",
                              maxIter = 200, msMaxIter = 200), method="REML")

anova(s.final)
```


Diagnostics for RE 


###Chinook



###Pink

###Chum

###Coho


#General Conclusions

