---
title: "Explore-Coastwide-Data"
author: "Dr. Curry J. Cunningham"
date: "2/5/2019"
output:
  html_document:
    highlight: espresso
    number_sections: yes
    smooth_scroll: yes
    theme: sandstone
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
require(tidyverse)
require(kableExtra)
require(dplyr)
require(ggthemes)
```

#Description of Data
* Salmon stock-recruitment data for populations in Alaska, BC, and Puget Sound were compiled from various sources by C. Cunningham (AFSC).
* Salmon SR data from Washington, Oregon, and California were compiled by J. Gosselin and B. Burke (NWFSC) 
* These two datasets were combined in late November, 2018. 
* **Update:** Chinook salmon data have multiple runs with the same `stock` identifier. Scripts must contain code to append `run` to `stock` if present. 

#Structure of Data
Data is now stored in object called  `AK-WCoast-Salmon-SR.csv`
```{r}
dat <- read.csv("AK-WCoast-Salmon-SR.csv", header=TRUE)
```

What is the structure of this data frame?
```{r}
kable(head(dat), align="crr") %>% #, digits=c(3,1,1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
Recorded brood years run from `r min(dat$broodYr)` to `r max(dat$broodYr)`.

#Preliminary Calculations
First we need to add recruits-per-spawner and log recruits-per-spawner
```{r}
dat$rps <- dat$rec/dat$spawn
dat$ln.rps <- log(dat$rps)
kable(head(dat), align="crr") %>% #, digits=c(3,1,1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

#Data Quality Control
We also need to screen out non-sensical values reported in the data, specifically:

* ln.rps is infinite, i.e. spawners were reported as zero
* ln.rps is `NA`, i.e. either `rec` or `spawn` was blank
* We also filter filter brood years: 

  + `broodYr > 1950` as this is the start of the NPGO index.
  + `broodYr <= 2010` to ensure all complete return of age classes for all species. 

```{r}
dat.2 <- dat %>% filter(!is.infinite(ln.rps), !is.na(ln.rps), broodYr>=1950, broodYr<=2010)
```

So, the total number of observations is: `r nrow(dat.2)`.

##Deal with multiple runs of Chinook salmon to same river
I noticed that on the West Coast there were multiple Chinook salmon `run` identifiers for the same `stock` by `broodYr` combination. We need to identify these instances and update the stock name so it is treated separately in subsequent data exploration and STAN stock-recruitment models. 

As it turns out there are `r sum(!is.na(dat.2$run))` observations (years) our of `r length(dat.2$run)` for which a `run` is identified.

To avoid confusion we will simply update `stock` names to include the `run` for each stock where it is listed. 

Lets determine for which `stock` identifier there are multiple years
```{r}
temp.dat <- dat.2 %>% filter(species=='Chinook')

stocks <- unique(temp.dat$stock)
n.stocks <- length(stocks)

n.runs <- vector(length=n.stocks)

for(s in 1:n.stocks) {
  n.runs[s] <- length(unique(temp.dat$run[temp.dat$stock==stocks[s]]))
}
stocks[which(n.runs>1)]
```

Looks like the Lewis River is our trouble spot.

So, lets append the run to the Lewis River in its `stock` designation. 

```{r}

```

#Explore Data
##Number of stocks by species and region
```{r}
sum.stocks_species.region <- dat.2 %>% group_by(species, region) %>% summarize('n.stocks'=length(unique(stock)))
# kable(sum.stocks_species.region, align="crr") %>% #, digits=c(3,1,1)) %>%
  # kable_styling(bootstrap_options = c("striped", "hover"))

g <- ggplot(sum.stocks_species.region, aes(x=region, y=n.stocks, fill=species)) +
       theme_linedraw() +
       scale_fill_colorblind() +
       geom_bar(stat='identity', position='stack') +
       theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
       ylab('Number of Stocks')
plot(g)
```

##Number of Years of Data by species and region
```{r}
sum.years_species.region <- dat.2 %>% group_by(species, region) %>% summarize('n.years'=n())
# kable(sum.years_species.region, align="crr") %>% #, digits=c(3,1,1)) %>%
  # kable_styling(bootstrap_options = c("striped", "hover"))

g <- ggplot(sum.years_species.region, aes(x=region, y=n.years, fill=species)) +
       theme_linedraw() +
       scale_fill_colorblind() +
       geom_bar(stat='identity', position='stack') +
       theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
       ylab('Number of Years with Data')
plot(g)
```

##Number of Regions is now: `r length(unique(dat.2$region))`.

#Plot Stock-Recruitment Data
##Species Combined
###Facet: species, Color: region
```{r, echo=FALSE, fig.height=8, fig.width=8}
g <- ggplot(dat.2, aes(x=broodYr, y=ln.rps, color=region, group=stock)) +
       theme_linedraw() +
       geom_line(alpha=0.25) +
       facet_wrap(~species, scales='free_y') +
       theme(legend.position='none')
plot(g)
```

###Facet: region, Color: species
```{r, echo=FALSE, fig.height=8, fig.width=8}
g <- ggplot(dat.2, aes(x=broodYr, y=ln.rps, color=species, group=stock)) +
       theme_linedraw() +
       scale_color_colorblind() +
       geom_line(alpha=0.25) +
       facet_wrap(~region, scales='free_y') +
       theme(legend.position='bottom')
plot(g)
```

##Chinook
###Data
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
temp.spec <- 'Chinook'
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=stock, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon'))
       # geom_smooth(aes(group=region))
plot(g)
```

###Data + Smooth
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=region, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon')) +
       geom_smooth(aes(group=region))
plot(g)
```

##Sockeye
###Data
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
temp.spec <- 'Sockeye'
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=stock, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon'))
       # geom_smooth(aes(group=region))
plot(g)
```

###Data + Smooth
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=region, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon')) +
       geom_smooth(aes(group=region))
plot(g)
```

##Pink
###Data
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
temp.spec <- 'Pink'
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=stock, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon'))
       # geom_smooth(aes(group=region))
plot(g)
```

###Data + Smooth
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=region, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon')) +
       geom_smooth(aes(group=region))
plot(g)
```

##Chum
###Data
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
temp.spec <- 'Chum'
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=stock, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon'))
       # geom_smooth(aes(group=region))
plot(g)
```

###Data + Smooth
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=region, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon')) +
       geom_smooth(aes(group=region))
plot(g)
```

##Coho
###Data
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
temp.spec <- 'Coho'
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=stock, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon'))
       # geom_smooth(aes(group=region))
plot(g)
```

###Data + Smooth
```{r, echo=FALSE, warn=FALSE, fig.height=8, fig.width=8}
g <- ggplot(filter(dat.2, species==temp.spec), aes(x=broodYr, y=ln.rps, color=region, group=factor(stock))) +
       geom_line(alpha=0.25) +
       facet_wrap(~region) +
       theme(legend.position='none') +
       ggtitle(paste(temp.spec,'Salmon')) +
       geom_smooth(aes(group=region))
plot(g)
```

